---
title: "Machine Learning - Learning Lab 1"
author: ""
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is an overview of the four labs in Machine Learning:

-   **Part 1: Create a clean data set for building logistic regression
    models.** We discuss a paper ([Predicting STEM and Non-STEM College
    Major Enrollment from Middle School Interaction with Mathematics
    Educational
    Software](https://educationaldatamining.org/EDM2014/uploads/procs2014/short%20papers/276_EDM-2014-Short.pdf))
    and the goal is to replicate the experiment in the paper. First, we
    need to understand the learning context, create a clean data set for
    the experiment, and visualize variables in the data set to gain an
    overview of it.
-   **Part 2: Build a baseline model.** We discuss the process of
    building a machine learning model. We'll learn one way of splitting
    and sampling data and then build a baseline model and evaluate model
    performance.
-   **Part 3: Cross validation.** We discuss another way of splitting
    and sampling data, cross validation.
-   **Part 4: Feature engineering.** We discuss the importance of
    feature engineering and how to communicate findings. We analyze
    confusion matrix, creatively add new features to improve model
    iteratively, and report findings.

In this learning lab, we will be focusing on Part 1, creating a clean
data set for building logistic regression models.

## Part 1.1. Understand the learning context

#### [Your Turn]{style="color: green;"} ⤵ {style="font-style: normal; font-variant-caps: normal; letter-spacing: normal; orphans: auto; text-align: start; text-indent: 0px; text-transform: none; white-space: normal; widows: auto; word-spacing: 0px; -webkit-tap-highlight-color: rgba(26, 26, 26, 0.3); -webkit-text-size-adjust: auto; -webkit-text-stroke-width: 0px; text-decoration: none; caret-color: rgb(0, 0, 0); color: rgb(0, 0, 0);"}

First, let's read [the brief (four page) paper that will help to set the
context for this learning
lab](Predicting%20STEM%20and%20Non-STEM%20College%20Major%20Enrollment%20from%20Middle%20School%20Interaction%20with%20Mathematics%20Educational%20Software).

Below, add notes on one or two of the following:

-   What kinds of learning activities are available in the ASSISTment
    system?
-   What's the research question?
-   What's the prediction task?
-   What are the variables used in the prediction task?
-   How might these variables be informative for the prediction task?
-   What kinds of new knowledge does this study add to the field of STEM
    education?

## Part 1.2. Access the data set

After reading the paper, we can see that the task is to predict whether
a student will choose STEM majors and the authors used the following
variables to predict STEM major enrollment: carelessness, knowledge,
correctness, boredom, engaged concentration, confusion, frustration,
off-task, gaming, and number of actions.

We will use a portion of the ASSISTment data from a [data mining
competition](https://sites.google.com/view/assistmentsdatamining/home?authuser=0)
to replicate the experiment in the paper, but the prediction task is not
whether a student will choose STEM majors, instead, it is predicting
whether the students (who have now finished college) pursue a career in
STEM fields (1) or not (0).

The data is in the folder `data`. There are multiple data sets in the
folder. We need to create one dataset from these data sets for building
models.

### Load the packages

We'll load three packages for this analysis.

```{r}
library(tidyverse)
library(here)
library(Hmisc) 
```

### Import the data

This will take little a while to load the data as these are big files.

```{r}
file_1 <- read_csv(here("data", "student_log_1.csv"))
file_2 <- read_csv(here("data", "student_log_5.csv"))
```

We'll merge these together using the `bind_rows()` function, which does
what it sounds like:

```{r}
dat <- bind_rows(file_1, file_2)
```

## Part 1.3. Create a clean data set

Look closely at the 11 data sets in the folder, you will find that files
starting with "student" do not have the prediction label (STEM or Not
STEM) and one student is in multiple rows in these files. We need to
delete duplicate rows in the files starting with "student" and combine
it with the file "prediction" (this file has the prediction label). For
more information about the meaning of columns, please read [this
file](https://docs.google.com/spreadsheets/d/1QVUStXiRerWbH1X0P11rJ5IsuU2Xutu60D1SjpmTMlk/edit?usp=sharing)
(very important!!!).

### Delete duplicate rows

ITEST_id is a deidentified ID/tag used for identifying an individual
student. Thus, we delete duplicate rows based on this column so that we
only have one row per student by using the `distinct()` function.

```{r}
dat_csv <- dat %>% 
  distinct(ITEST_id, .keep_all = TRUE)
```

#### [Your Turn]{style="color: green;"} ⤵ {style="font-style: normal; font-variant-caps: normal; letter-spacing: normal; orphans: auto; text-align: start; text-indent: 0px; text-transform: none; white-space: normal; widows: auto; word-spacing: 0px; -webkit-tap-highlight-color: rgba(26, 26, 26, 0.3); -webkit-text-size-adjust: auto; -webkit-text-stroke-width: 0px; text-decoration: none; caret-color: rgb(0, 0, 0); color: rgb(0, 0, 0);"}

Answer the two questions below by creating your own code chunk (by
clicking the green C in the upper right part of this pane) or by running
code below in the console.

-   How many students and variables does this data set (`dat_csv`) have?
-   What are the variables (from `dat_csv`) used in the experiment of
    the paper?

### Combine features for prediction and predicted label into one data set

Now, we have `dat_csv`, which is the data set containing students'
information. Next, let's load the data set containing predicted label.

```{r}
dat_csv_predict <- read_csv(here("data", "prediction.csv"))
```

#### [Your Turn]{style="color: green;"} ⤵ {style="font-style: normal; font-variant-caps: normal; letter-spacing: normal; orphans: auto; text-align: start; text-indent: 0px; text-transform: none; white-space: normal; widows: auto; word-spacing: 0px; -webkit-tap-highlight-color: rgba(26, 26, 26, 0.3); -webkit-text-size-adjust: auto; -webkit-text-stroke-width: 0px; text-decoration: none; caret-color: rgb(0, 0, 0); color: rgb(0, 0, 0);"}

Like earlier, answer the questions below by creating your own code chunk
(by clicking the green C in the upper right part of this pane) or by
running code below in the console.

-   How many students and variables does this data set
    (`dat_csv_predict`) have?
-   What is the predicted label (from `dat_csv_predict`)?

Next, let's combine `dat_csv` and `dat_csv_predict`. You might have
noticed that we will get less than 1000 instances for the experiment
after combining the two data sets. This is a very small data set. For
this lab, we use this small data set for practice. In your own project,
it would be good if you have more than 1,000 instances and the more, the
better.

We will use `ITEST_id` to combine these two data sets.

```{r}
dat_csv_combine <- inner_join(dat_csv, dat_csv_predict, by = "ITEST_id")
```

#### [Your Turn]{style="color: green;"} ⤵ {style="font-style: normal; font-variant-caps: normal; letter-spacing: normal; orphans: auto; text-align: start; text-indent: 0px; text-transform: none; white-space: normal; widows: auto; word-spacing: 0px; -webkit-tap-highlight-color: rgba(26, 26, 26, 0.3); -webkit-text-size-adjust: auto; -webkit-text-stroke-width: 0px; text-decoration: none; caret-color: rgb(0, 0, 0); color: rgb(0, 0, 0);"}

-   How many students and variables does this combined dataset have?
-   What are the variables used in the experiment of the paper?
-   What is the predicted label?

Next, let's keep only the variables that we will use for the experiment.

```{r}
dat_csv_combine <- dat_csv_combine %>% 
  select(AveCarelessness, AveKnow, AveCorrect.x, AveResBored, 
         AveResEngcon, AveResConf, AveResFrust, AveResOfftask, 
         AveResGaming, NumActions, isSTEM)
```

Let's save the data set into a CSV file so that we can use it in the
following learning labs in this module.

```{r}
write_csv(dat_csv_combine, here("data", "dat_csv_combine.csv"))
```

## Part 1.4. Visualize variables in the combined dataset

Now, we get the data set for modeling. Before modeling, let's look at
the relationship between features for prediction and predicted label one
by one and evaluate whether these features would be informative for the
prediction. Let's start with the relationship between `AveCarelessness`
and `isSTEM`.

### Create a visualization with two variables

Create a bar with mean of `AveCarelessness` and `isSTEM`.

```{r}
ggplot(dat_csv_combine, aes(x = isSTEM, y = AveResOfftask)) +
  stat_summary(fun.data = mean_sdl, geom = "bar")
```

#### [Your Turn]{style="color: green;"} ⤵ {style="font-style: normal; font-variant-caps: normal; letter-spacing: normal; orphans: auto; text-align: start; text-indent: 0px; text-transform: none; white-space: normal; widows: auto; word-spacing: 0px; -webkit-tap-highlight-color: rgba(26, 26, 26, 0.3); -webkit-text-size-adjust: auto; -webkit-text-stroke-width: 0px; text-decoration: none; caret-color: rgb(0, 0, 0); color: rgb(0, 0, 0);"}

-   What do you see in this bar chart?

Now create more bars to explore the relationship between features (i.e.,
carelessness, knowledge, correctness, boredom, engaged concentration,
confusion, frustration, off-task, gaming, and number of actions) for
prediction and predicted label.

#### [Your Turn]{style="color: green;"} ⤵ {style="font-style: normal; font-variant-caps: normal; letter-spacing: normal; orphans: auto; text-align: start; text-indent: 0px; text-transform: none; white-space: normal; widows: auto; word-spacing: 0px; -webkit-tap-highlight-color: rgba(26, 26, 26, 0.3); -webkit-text-size-adjust: auto; -webkit-text-stroke-width: 0px; text-decoration: none; caret-color: rgb(0, 0, 0); color: rgb(0, 0, 0);"}

```{r}
#write your code here

```

-   What do you see in the bar charts?

### Create a visualization with three variables

Now create a scatter plot with two variables from the features for
prediction and the predicted label. In the scatter plot, the two
variables will be x or y axis and the predicted label will be the color
of dots.

```{r}
#write your code here
ggplot(dat_csv_combine, aes(x = AveResOfftask, y = AveKnow, color = isSTEM)) +
  geom_point()
```

-   What do you see in the scatter plots?

## Part 1.5. Communicate your insights

-   How might the selected features be informative for the prediction
    task?

## Reach

For this learning lab, your reach is to create your own dataset. This
dataset should include features for prediction and predicted label.
Before creating such a dataset, you should consider what your prediction
task is (determining what the predicted label will be) and what might be
related to the predicted label (determining what features will be
informative). This requires in-depth domain knowledge, which you are an
expert on!
